---
title: "1_clean"
author: "Thomas E. Metherell"
date: "2022-12-06"
output: html_document
---

Script in series: "Associations between early life mental health indicators and sleep in adulthood"

This script cleans and pre-processes main survey data from the 1970 British Cohort Study, along with selected CLOSER harmonisation work packages, for use in further analysis.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dependencies
We start by loading the necessary dependencies.

```{r dependencies}
library(haven)
library(tidyverse)
library(magrittr)
```

## Loading data
We can now load the raw BCS70 and CLOSER data downloaded from the UK Data Service. Data in this project are handled in Stata (.dta) format.

```{r load_data}
# Reading DTA files containing BCS70 data from sweeps at birth, ages 5 (x3), 10, 16 (x4), 26, 30, 42 and 46
# Storing the contents of each as a separate data frame
birth <- read_dta("..\\..\\Data\\Raw\\1a_Birth_&_22_month\\UKDA-2666-stata9\\stata9\\bcs7072a.dta")
age_5e <- read_dta("..\\..\\Data\\Raw\\2_age_5\\UKDA-2699-stata\\stata\\stata11\\f699b.dta")
age_5f <- read_dta("..\\..\\Data\\Raw\\2_age_5\\UKDA-2699-stata\\stata\\stata11\\f699c.dta")
age_5dv <- read_dta("..\\..\\Data\\Raw\\2_age_5\\UKDA-2699-stata\\stata\\stata11\\bcs2derived.dta")
age_10 <- read_dta("..\\..\\Data\\Raw\\3_age_10\\UKDA-3723-stata\\stata\\stata11_se\\sn3723.dta")
age_16 <- read_dta("..\\..\\Data\\Raw\\4_age_16\\UKDA-3535-stata\\stata\\stata13_se\\bcs7016x.dta")
age_16arith <- read_dta("..\\..\\Data\\Raw\\4a_age_16_arithmetic\\UKDA-6095-stata\\stata\\stata11\\bcs70_16-year_arithmetic_data.dta")
age_16dv <- read_dta("..\\..\\Data\\Raw\\4_age_16\\UKDA-3535-stata\\stata\\stata13_se\\bcs4derived.dta")
age_16school <- read_dta("..\\..\\Data\\Raw\\4_age_16\\UKDA-3535-stata\\stata\\stata13_se\\bcs70_age16_school_type.dta")
age_26 <- read_dta("..\\..\\Data\\Raw\\5_age_26\\UKDA-3833-stata\\stata\\stata11\\bcs96x.dta")
age_30 <- read_dta("..\\..\\Data\\Raw\\6_age_30\\UKDA-5558-stata\\stata\\stata11_se\\bcs2000.dta")
age_42 <- read_dta("..\\..\\Data\\Raw\\9_age_42\\UKDA-7473-stata\\stata\\stata13\\bcs70_2012_flatfile.dta")
age_46 <- read_dta("..\\..\\Data\\Raw\\10_age_46\\UKDA-8547-stata\\stata\\stata11\\bcs_age46_main.dta")

# Reading DTA files containing data from CLOSER WPs 1, 2, 9 and 20
# Storing the contents of each as a separate data frame
CLOSER_MH <- read_dta("..\\..\\Data\\Raw\\CLOSER_age_10_mental_health\\UKDA-8882-stata\\stata\\stata13\\bcs70_closer_wp20.dta")
CLOSER_env <- read_dta("..\\..\\Data\\Raw\\CLOSER_childhood_environment_adult_wellbeing\\UKDA-8553-stata\\stata\\stata13\\bcs70_closer_wp9.dta")
CLOSER_BMI <- read_dta("..\\..\\Data\\Raw\\CLOSER_height_weight_BMI\\UKDA-8551-stata\\stata\\stata13\\bcs70_closer_wp1.dta")
CLOSER_SES <- read_dta("..\\..\\Data\\Raw\\CLOSER_socioeconomic_measures\\UKDA-8305-stata\\stata\\stata13\\bcs70_closer_wp2ses.dta")

# Reading DTA file of response dataset and storing the contents as a data frame
response <- read_dta("..\\..\\Data\\Raw\\Response_dataset\\UKDA-5641-stata\\stata\\stata13\\bcs70_response_1970-2016.dta")
```

## Selecting relevant variables
We now select the variables relevant to our analyses (see study protocol), discarding the unnecessary data.

```{r select_vars}
# Birth sweep data
birth %<>% select(
  bcsid, # BCS70 serial number
  a0014, # Social class of father in 1970* 
  a0018, # Mother's social class in 1970*
  a0230, # Occurrence of genital tract bleeding*
  a0255 # Sex of the baby*
)

# Age 5 sweep data (E)
age_5e %<>% select(
  bcsid, # BCS70 serial number
  e192 # Age father left school*
)

# Age 5 sweep data (F)
age_5f %<>% select(
  bcsid, # BCS70 serial number
  f100, # Schonell Reading Test score*
  f118, # Complete a Profile Test score*
  f121, # Standardised Human Figure Drawing score*
  f122 # Standardised Copying Designs score*
)

# Renaming "BCSID" to "bcsid" in age_5dv for consistency
age_5dv %<>% rename(bcsid = BCSID)

# Age 5 derived variables
age_5dv %<>% select(
  bcsid, # BCS70 serial number
  BD2READ # Standardised English Picture Vocabulary Test score*
)

# Age 10 sweep data
age_10 %<>% select(
  bcsid, # BCS70 serial number
  c3_11, # Mother's corrected social class
  c9_1:c9_7, # Total family income
  i3504:i3540, # BAS word definitions
  i3541:i3574, # BAS recall of digits
  i3617:i3644, # BAS matrices
  i4201:i4221, # BAS word similarities (derived)
  m43:m61, # Rutter scale
  m84:m101, # Positive activities outside school*
  j127:j178b, # Child Development Scale
  tests10 # Maths test, reading test, BAS test completed
)

# Age 16 sweep data
age_16 %<>% select(
  bcsid, # BCS70 serial number
  contains("hd5_"), # Alcohol in last 7 days*
  contains("oe1_"), # Sources of income*
  rd6m_1 # Behavioural or emotional problems*
)

# Age 16 arithmetic test data
age_16arith %<>% select(
  bcsid, # BCS70 serial number
  mathscore # Arithmetic test score*
)

# Renaming "BCSID" to "bcsid" in age_16dv for consistency
age_16dv %<>% rename(bcsid = BCSID)

# Age 16 derived variables
age_16dv %<>% select(
  bcsid, # BCS70 serial number
  BD4MAL, # Total Malaise score
  BD4RUTT # Total Rutter behaviour score
)

# Renaming "BCSID" to "bcsid" in age_16school for consistency
age_16school %<>% rename(bcsid = BCSID)

# Age 26 sweep data
age_26 %<>% select(
  bcsid, # BCS70 serial number
  b960540 # Accidents and assaults*
)

# Age 30 sweep data
age_30 %<>% select(
  bcsid, # BCS70 serial number
  aptrain, # Apprenticeships*
  emosup, # People for advice/support*
  numadmn, # Hospital admissions*
  wrktrain, # Work-related training*
  vote97 # Voted in 1997 General Election*
)

# Renaming "BCSID" to "bcsid" in age_42 for consistency
age_42 %<>% rename(bcsid = BCSID)

# Age 42 sweep data
age_42 %<>% select(
  bcsid, # BCS70 serial number
  B9DIVCHK, B9HMS, B9MARCHK, # Marital status*
  contains("B9SCQ8"), # Organisational membership*
  B9SCQ17, # Computer ownership*
  B9SCQ19, # Number of cars or vans owned*
  B9TEN # Housing tenure*
)

# Renaming "BCSID" to "bcsid" in age_46 for consistency
age_46 %<>% rename(bcsid = BCSID)

# Age 46 sweep data
age_46 %<>% select(
  bcsid, # BCS70 serial number
  BD10MAL, # Total Malaise score
  contains("B10MHPRB"), # Seen Dr/specialist for mental health problems
  BD10WEMWB # Warwick Edinburgh Mental Well-Being Scale
)

# CLOSER WP9 data
CLOSER_env %<>% select(
  bcsid, # BCS70 serial number
  tenure, # Accommodation owned or rented in childhood
  prmnh, # Poor maternal / familial mental health
  intbcsz, # Internalising behaviour in childhood, standardised
  extbcsz # Externalising behaviour in childhood, standardised
)

# CLOSER WP20 data
CLOSER_MH %<>% select(
  -closerid, -stid # Selecting all variables except CLOSER and study identifiers
)

# Restricting CLOSER WP1 data to those collected at age 10 only
CLOSER_BMI <- CLOSER_BMI[CLOSER_BMI$visitage == 10,]

# CLOSER WP1 data
CLOSER_BMI %<>% select(
  bcsid, # BCS70 serial number
  bmi # Body mass index
)

# CLOSER WP2 data
CLOSER_SES %<>% select(
  bcsid, # BCS70 serial number
  fclrg90 # Father's social class (RG 1990 version) at age 10 sweep
)

# Renaming "BCSID" to "bcsid" in response for consistency
response %<>% rename(bcsid = BCSID)

# Response dataset
response %<>% select(
  bcsid, # BCS70 serial number
  contains("OUTCME") # Outcome variables
)
```

## Defining NAs
In order to allow correct data analysis, we define data points which are not available as NA. Mostly these are negative values from the source datasets, but there are some exceptions.

```{r define_na}
# Birth dataset: negative values become NA
is.na(birth[,]) <- birth[,] < 0

# Age 5 dataset (E): negative values become NA
is.na(age_5e[,]) <- age_5e[,] < 0

# Age 5 dataset (F, reading & profile tests): negative values become NA
is.na(age_5f[c("f100", "f118")]) <- age_5f[c("f100", "f118")] < 0

# Age 5 dataset (F, HFD & copying designs tests): -5 becomes NA
is.na(age_5f[c("f121", "f122")]) <- age_5f[c("f121", "f122")] == -5

# Age 5 dataset (derived): -1 becomes NA
is.na(age_5dv[,]) <- age_5dv[,] == -1

# Age 10 dataset (mother's social class only): negative values and values more than 7 become NA
is.na(age_10[, "c3_11"]) <- age_10[, "c3_11"] < 0
is.na(age_10[, "c3_11"]) <- age_10[, "c3_11"] > 7

# Age 10 dataset (positive activities): negative values become NA
is.na(age_10[, grep("m84", names(age_10)):grep("m101", names(age_10))]) <- age_10[, grep("m84", names(age_10)):grep("m101", names(age_10))] < 0

# Age 16 dataset: negative values become NA
is.na(age_16[,]) <- age_16[,] < 0

# Age 16 derived variables dataset: negative values become NA
is.na(age_16dv[,]) <- age_16dv[,] < 0

# Age 16 school type dataset: negative values become NA
is.na(age_16school[,]) <- age_16school[,] < 0

# Age 26 dataset: negative values become NA
is.na(age_26[,]) <- age_26[,] < 0

# Age 30 dataset (hospital admissions only): 99 becomes NA
is.na(age_30[, "numadmn"]) <- age_30[, "numadmn"] == 99

# Age 30 dataset (all other variables): values above 7 become NA
is.na(age_30[c("aptrain", "wrktrain", "emosup", "vote97")]) <- age_30[c("aptrain", "wrktrain", "emosup", "vote97")] > 7

# Age 42 dataset: negative values become NA
is.na(age_42[,]) <- age_42[,] < 0

# Age 46 dataset: negative values become NA
is.na(age_46[,]) <- age_46[,] < 0

# CLOSER SES dataset: negative values become NA
is.na(CLOSER_SES[,]) <- CLOSER_SES[,] < 0

# CLOSER childhood environment dataset (all variables except internalising/externalising behaviour): negative values become NA
is.na(CLOSER_env[c("prmnh", "tenure")]) <- CLOSER_env[c("prmnh", "tenure")] < 0

# CLOSER childhood environment dataset (internalising/externalising behaviour): values less than -100 become NA
is.na(CLOSER_env[c("intbcsz", "extbcsz")]) <- CLOSER_env[c("intbcsz", "extbcsz")] < -100
```

## Cognitive ability - principal component analysis
In order to use cognitive ability as an exposure/confounder in our study, we need to convert the raw cognitive test scores into a univariate measure of cognitive ability. We do this by performing a principal component analysis (PCA), with the first component being the cognitive ability measure used in the study. This is performed at both age 5 and age 10.

```{r cognitive_ability_pca}
## Age 5
# Joining derived dataset (containing BD2READ) into dataset F and deleting the former
age_5f %<>% full_join(., age_5dv, by = "bcsid")
rm(age_5dv)

# Detecting missing values in rows
missing <- rowSums(is.na(age_5f)) != 0
age_5f <- data.frame(age_5f, missing)

# Performing principal component analysis
pca <- prcomp(age_5f[age_5f$missing == FALSE, c("BD2READ", "f100", "f118", "f121", "f122")], center = TRUE, scale. = TRUE)

# Initialising derived cognitive ability score variable
age_5f$dv_cog_abil_5 <- rep(NA, nrow(age_5f))

# Storing PCA first component as derived cognitive ability score
age_5f[age_5f$missing == FALSE,]$dv_cog_abil_5 <- pca[["x"]][, 1]

# Scaling the derived score to have mean 0 and SD 1
age_5f$dv_cog_abil_5 %<>% scale(.) %>% as.numeric(.)

# Removing cognitive ability other than the derived score
age_5f %<>% select(-BD2READ, -f100, -f118, -f121, -f122, -missing)

## Age 10
# Summing scores from the four BAS tests conducted, or reporting NA if the BAS were not completed
age_10 %<>% mutate(
  
  # Word Definitions
  dv_bas_worddef = case_when(
    tests10 == 1 ~ rowSums(across(i3504:i3540, `==`, 1)),
    tests10 == 0 ~ NA_real_
  ),
  
  # Recall of Digits
  dv_bas_recall = case_when(
    tests10 == 1 ~ rowSums(across(i3541:i3574, `==`, 1)),
    tests10 == 0 ~ NA_real_
  ),
  
  # Matrices
  dv_bas_matrices = case_when(
    tests10 == 1 ~ rowSums(across(i3617:i3644, `==`, 1)),
    tests10 == 0 ~ NA_real_
  ),
  
  # Word Similarities
  dv_bas_wordsim = case_when(
    tests10 == 1 ~ rowSums(across(i4201:i4221, `==`, 1)),
    tests10 == 0 ~ NA_real_
  )
)

# Performing principal component analysis
pca <- prcomp(age_10[age_10$tests10 == 1,] %>% select(contains("dv_bas")), center = TRUE, scale. = TRUE)

# Initialising derived cognitive ability score variable
age_10$dv_bas_g <- rep(NA, nrow(age_10))

# Storing PCA first component as derived cognitive ability score
age_10[age_10$tests10 == 1,]$dv_bas_g <- -pca[["x"]][, 1]

# Scaling the derived score to have mean 0 and SD 1
age_10$dv_bas_g %<>% scale(.) %>% as.numeric(.)

# Removing cognitive ability other than the derived score
age_10 %<>% select(
  -(contains("dv_bas") & !dv_bas_g),
  -(i3504:i4221),
  -tests10
)
```

## Merging the datasets
We can now merge the individual datasets into a single 'wide' dataset with one row per participant.

```{r merge_datasets}
# Initialising final dataset by copying over birth data
data_clean <- birth

# Joining each of the other datasets in turn to the final dataset using the BCS70 serial numbers of the participants
for(dataset in c("age_5e", "age_5f", "age_10", "age_16", "age_16arith", "age_16dv", "age_16school", "age_26", "age_30", "age_42", "age_46", "CLOSER_MH", "CLOSER_env", "CLOSER_BMI", "CLOSER_SES", "response")){
  data_clean <- full_join(data_clean, get(dataset), by = "bcsid")
}
```

## Deriving control variables
To make the data useful, we produce a number of derived variables which are more informative than the raw source data.

```{r control_variables}
# Deriving maximum parental social class at birth and removing mother variable
data_clean %<>% rowwise() %>% mutate(dv_sc_birth = min(a0014, a0018))
data_clean %<>% select(-a0018)

# Recoding "other"/"unsupported" values as missing in a0014 for use as an auxiliary variable
is.na(data_clean[c("a0014")]) <- data_clean[c("a0014")] > 6

# Recoding father's education as a binary factor variable for use as an auxiliary variable
data_clean %<>% mutate(dv_father_schl = case_when(
  e192 >= 16 ~ 1,
  e192 < 15 ~ 0
))
data_clean$dv_father_schl %<>% factor(.)
data_clean %<>% select(-e192)

# Combining age 10 income variables into a single variable and removing individual variables
data_clean %<>% mutate(dv_income_10 = case_when(
  c9_1 == 1 ~ 1,
  c9_2 == 1 ~ 2,
  c9_3 == 1 ~ 3,
  c9_4 == 1 ~ 4,
  c9_5 == 1 ~ 5,
  c9_6 == 1 ~ 6,
  c9_7 == 1 ~ 7
))
data_clean %<>% select(-contains("c9"))

# Deriving maximum parental social class at age 10 and removing individual variables
data_clean %<>% rowwise() %>% mutate(dv_sc_age_10 = min(fclrg90, c3_11))
data_clean %<>% select(-fclrg90, -c3_11)

# Combining age 16 income source variables into a single variable (employment & investments vs other) and removing individual variables
data_clean %<>% mutate(dv_incsource = case_when(
  if_any(oe1_1:oe1_5, ~. == 1) ~ 1,
  if_all(oe1_1:oe1_5, ~. == 2) & if_any(oe1_6:oe1_19, ~. == 1) ~ 2
))
data_clean$dv_incsource %<>% factor(.)
data_clean %<>% select(-contains("oe1_"))

# Combining positive activities outside school into a single variable and removing individual variables
data_clean %<>% mutate(dv_pos_act = rowSums(across(m84:m101)))
data_clean %<>% select(-(m84:m101))

# Combining alcohol consumption in the last week into a single variable and removing individual variables
data_clean %<>% mutate(dv_alcohol = case_when(
  if_any(hd5_1:hd5_7, ~. == 1) ~ 1,
  hd5_8 == 1 ~ 0
))
data_clean$dv_alcohol %<>% factor(.)
data_clean %<>% select(-contains("hd5_"))

# Combining apprenticeships and work-related training into a single variable and removing individual variables
data_clean %<>% mutate(dv_training = case_when(
  if_any(c("aptrain", "wrktrain"), ~. == 1) ~ 1,
  if_all(c("aptrain", "wrktrain"), ~. == 2) ~ 0
))
data_clean$dv_training %<>% factor(.)
data_clean %<>% select(-aptrain, -wrktrain)

# Recoding car ownership as a binary factor variable
data_clean %<>% mutate(B9SCQ19 = case_when(
  B9SCQ19 > 0 ~ 1,
  B9SCQ19 == 0 ~ 0
))
data_clean$B9SCQ19 %<>% factor(.)

# Combining response outcome variables into a single binary variable and removing individual variables
data_clean %<>% mutate(dv_participation = case_when(
  if_all(contains("OUTCME"), ~. == 1) ~ 1,
  if_any(contains("OUTCME"), ~. != 1) ~ 0
))
data_clean$dv_participation %<>% factor(.)
data_clean %<>% select(-contains("OUTCME"))

# Combining organisation membership into a single variable and removing individual variables
data_clean %<>% mutate(dv_organisations = case_when(
  if_any(B9SCQ8A:B9SCQ8P, ~. == 1) ~ 1,
  if_all(B9SCQ8A:B9SCQ8P, ~. == 0) ~ 0
))
data_clean$dv_organisations %<>% factor(.)
data_clean %<>% select(-(B9SCQ8A:B9SCQ8P))

# Combining marital status items into a single variable and removing individual variables
data_clean %<>% mutate(dv_marital = case_when(
  B9MARCHK == 1 | B9HMS %in% c(2, 5) ~ "Married/CP",
  B9DIVCHK == 1 | B9HMS %in% c(1, 3, 4, 6, 7) ~ "Divorced/Separated/Widowed",
  B9HMS == 8 ~ "Single"
))
data_clean %<>% select(-B9DIVCHK, -B9HMS, -B9MARCHK)

# Recoding housing tenure at age 42
data_clean %<>% mutate(B9TEN = case_when(
  B9TEN %in% c(1, 2) ~ "Owned",
  B9TEN %in% c(3, 4, 5, 7) ~ "Not owned"
))

# Deriving total Rutter score at age 10
data_clean %<>% mutate(dv_rutter_10 = rowSums(across(m43:m61)))
data_clean %<>% select(-(m43:m61))

# Deriving total CDS score at age 10
data_clean %<>% mutate(j178a = 48 - j178a, j178b = 48 - j178b)
data_clean %<>% mutate(dv_cds_10 = rowSums(across(j127:j178b)))
data_clean %<>% select(-(j127:j178b))
```

## Exporting data
Finally, we can export the cleaned dataset for future use.

```{r export_clean_data}
write_dta(data_clean, ".\\Cleaned\\data_clean.dta")
```
