---
title: "2d_winkler"
author: "Thomas E. Metherell"
date: "2023-02-08"
output: html_document
---

Script in series: "Associations between early life mental health indicators and sleep in adulthood"

This script is a naïve R implementation of Danielle Bodicoat's (07/10/2015) Stata do-file implementing Winkler et al.'s (2016) algorithm for deriving sleep time from raw ActivPAL event files.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dependencies
We start by loading the necessary dependencies.

```{r dependencies}
library(parallel)
library(purrr)
library(data.table)
```

## Splitting data
Because this algorithm is computationally intensive, we will use parallel computation to speed up the process. To do this we need to divide up the participants into as many groups as there are cores available for use.

```{r split}
# Setting random seed
set.seed(3154)

# Creating a list of event files
filelist <- list.files(path = ".\\Events_EventMarker", pattern = "Events.csv")

# Generating a random integer between 1 and the number of cores available for each participant, then creating a data frame of them
node <- sample(1:detectCores(), length(filelist), replace = TRUE)
assignments <- data.frame(1:length(node), node)

# Splitting the file list according to the random number assigned
for(i in 1:detectCores()){
  assign(
    paste0("filelist", i),
    filelist[assignments$node == i]
  )
}

# Removing unnecessary objects
rm(list = c("filelist", "node", "assignments", "i"))
```

## The algorithm
We can now implement the Winkler et al. (2016) algorithm as detailed below. The program iterates through participants and successively appends the results to a .csv file; make sure no .csv file of the same name already exists before running!

```{r processing}
# Establishing parallel computation cluster
## NB: THIS SECTION IS CURRENTLY SET TO CREATE AS MANY NODES AS THERE ARE CPU CORES AVAILABLE. THIS MAY CAUSE PROBLEMS IF THERE ARE OTHER TASKS RUNNING SIMULTANEOUSLY
### if you need to reduce the number, substitute `detectCores()` with e.g. `detectCores() - 2` or `6` in BOTH of the appearances below AND the one above in the data splitting section
cl <- makeCluster(detectCores())

# Loading dependencies within the cluster environment
clusterEvalQ(cl, {
  library(data.table)
  library(tidyverse)
  library(magrittr)
  library(lubridate)
  library(chron)
})

# Making the filelists available in the cluster environment
clusterExport(cl, ls(pattern = "filelist"))

# Applying the algorithm in the cluster
parLapply(cl = cl, X = 1:detectCores(), fun = function(no){
  
  # Initialising data frames
  assign(paste0("df", no),
         data.frame(ParticipantId = as.integer(), md_winkler_sleepdur = as.double()))
  
  # Looping through participants
  for(i in get(paste0("filelist", no))){
    
    # Reading the participant's *Events.csv file
    df <- read.csv(paste0(".\\Events_EventMarker\\", i))
    
    df %<>%
      # Giving variables more sensible names
      rename(
        Intervals = "Interval..s.",
        ActivityType = "ActivityCode..0.sedentary..1..standing..2.stepping."
      ) %>%
      
      # Retaining only needed variables
      select(Time, Intervals, ActivityType, CumulativeStepCount) %>%
      
      # Removing duplicate rows
      distinct() %>%
      
      mutate(
        # Creating variable encoding number of steps in current bout
        steps = if_else(
          row_number() == 1,
          CumulativeStepCount,
          CumulativeStepCount - lag(CumulativeStepCount)
        ),
        
        # Marking bouts where the activity type is different from that of the previous bout
        changeofactivity = if_else(
          row_number() == 1 | ActivityType != lag(ActivityType),
          TRUE,
          FALSE
        ),
        
        # Creating an 'event ID', i.e. running count of spells of a single activity type, derived from the cumulative sum of changes of activity type
        event = cumsum(changeofactivity)
      ) %>%
      
      # Removing unnecessary variables
      select(-CumulativeStepCount, -changeofactivity) %>%
      
      # Grouping by event ID
      group_by(event) %>%
      
      mutate(
        # Replacing bout duration with the duration of the entire event
        Intervals = sum(Intervals),
        
        # Replacing step count with the sum of steps in the entire event
        steps = sum(steps),
        
        # Replacing bout start time with the start time of the entire event
        Time = Time[row_number() == 1]
      ) %>%
      
      # Removing duplicate rows so there is only one row per event
      distinct() %>%
      
      # Ungrouping
      ungroup() %>%
      
      mutate(
        # Formatting date/time into standard POSIXct format
        Time = strptime("1899-12-31 00:00:00", format = "%Y-%m-%d %H:%M:%S") + Time * 86400,
        
        # Separately encoding the date and time
        date = as_date(strftime(Time, format = "%Y-%m-%d")),
        timeofday = times(strftime(Time, format = "%H:%M:%S")),
        
        # Creating a 'reference day' variable (noon to noon), where times in the morning are associated with the previous day
        refday = case_when(
          timeofday >= times("12:00:00") ~ date,
          timeofday < times("12:00:00") ~ date - 1
        )) %>%
      
      # Removing unnecessary variables
      select(-date, -timeofday) %>%
      
      # Grouping by reference day
      group_by(refday) %>%
      
      # Sorting by time within groups
      arrange(Time, .by_group = TRUE) %>%
      
      mutate(
        # Creating a new binary variable indicating whether an event is counted as sleep, and initially classing any sedentary/standing event over 5 hours, or the longest sedentary/standing event (if it exceeds 2 hours), as sleep
        sleep = case_when(
          ActivityType %in% c(0,1) & Intervals >= 18000 ~ TRUE,
          Intervals == max(Intervals[ActivityType %in% c(0, 1)]) & Intervals >= 7200 ~ TRUE
        ),
        sleep = if_else(
          is.na(sleep),
          FALSE,
          sleep
        )
      )
    
    # Creating dummy variables for iteration
    old <- -1
    new <- 0
    
    # Iterating until the number of rows classed as 'sleep' stops increasing
    while(old < new){
      
      # Setting initial 'sleep' row count
      old <- new
      
      df %<>%
        mutate(
          # Creating a variable 'sleep_end' that encodes the end of the 'sleep window' for events currently marked as sleep
          sleep_end = if_else(
            sleep,
            Time + Intervals + 900,
            NA_POSIXct_
          ),
          
          # If the current event is within the previous event's sleep window, extending sleep_end to the current event
          sleep_end = if_else(
            is.na(sleep_end) & Time < lag(sleep_end) & !is.na(lag(sleep_end)),
            lag(sleep_end),
            NA_POSIXct_
          ),
          
          # If an event is the last within a sleep window, and is sedentary with a duration of >= 2h, marking it as sleep
          sleep = if_else(
            !is.na(sleep_end) & is.na(lead(sleep_end)) & ActivityType %in% c(0, 1) & Intervals >= 7200,
            TRUE,
            sleep
          )
        ) %>%
        
        # Regrouping by reference day and sleep_end value (i.e. sleep window)
        ungroup() %>% group_by(refday, sleep_end) %>%
        
        # Summing number of steps within sleep windows
        mutate(total_end = if_else(
          !is.na(sleep_end),
          sum(steps),
          NA_integer_
        )) %>%
        
        # Regrouping by reference day only
        ungroup() %>% group_by(refday) %>%
        
        mutate(
          # If an event is the last within a sleep window, and there have been less than 20 steps within that window, and either this or the following event is sedentary with a duration of >= 30 minutes, marking it as sleep
          sleep = if_else(
            !is.na(sleep_end) & is.na(lead(sleep_end)) & total_end <= 20 & ((ActivityType %in% c(0, 1) & Intervals >= 1800) | (lead(ActivityType) %in% c(0, 1) & lead(Intervals) >= 1800)),
            TRUE,
            sleep
          ),
          
          # If the event following this event is marked as sleep, and the event is within a sleep window, marking it as sleep
          sleep = if_else(
            lead(sleep) & !is.na(sleep_end),
            TRUE,
            sleep
          )
        ) %>%
        
        # Sorting in reverse time order by group
        ## NB: this means that lag() = the next event, lead() = the previous event
        arrange(desc(Time), .by_group = TRUE) %>%
        
        mutate(
          # Creating a variable 'sleep_start' that encodes the end of the 'sleep window' for events currently marked as the beginning of a series of sleep events
          sleep_start = if_else(
            is.na(lead(sleep)) & sleep,
            Time - 900,
            NA_POSIXct_
          ),
          
          # If the current event is within the next event's sleep window, extending sleep_start to the current event
          sleep_start = if_else(
            is.na(sleep_start) & (Time + Intervals) > lag(sleep_start) & !is.na(lag(sleep_start)),
            lag(sleep_start),
            sleep_start
          ),
          
          # If an event is the first within a sleep window, and is sedentary with a duration of >= 2h, marking it as sleep
          sleep = if_else(
            !is.na(sleep_start) & is.na(lead(sleep_start)) & ActivityType %in% c(0, 1) & Intervals >= 7200,
            TRUE,
            sleep
          )
        ) %>%
        
        # Sorting in normal time order by group
        ## NB: lag() and lead() return to functioning as would be expected
        arrange(Time, .by_group = TRUE) %>%
        
        # Regrouping by reference day and sleep_start value
        ungroup() %>% group_by(refday, sleep_start) %>%
        
        # Summing number of steps within sleep windows
        mutate(total_start = if_else(
          !is.na(sleep_start),
          sum(steps),
          NA_integer_
        )) %>%
        
        # Regrouping by reference day only
        ungroup() %>% group_by(refday) %>%
        
        mutate(
          # If an event is the first within a sleep window, and there will be less than 20 steps within that window, and either this or the preceding event is sedentary with a duration of >= 30 minutes, marking it as sleep
          sleep = if_else(
            !is.na(sleep_start) & is.na(lag(sleep_start)) & total_start <= 20 & ((ActivityType %in% c(0, 1) & Intervals >= 1800) | (lag(ActivityType) %in% c(0, 1) & lag(Intervals) >= 1800)),
            TRUE,
            sleep
          ),
          
          # If the event preceding this event is marked as sleep, and the event is within a sleep window, marking it as sleep
          sleep = if_else(
            lag(sleep) & !is.na(sleep_start),
            TRUE,
            sleep
          ),
          
          # Marking as sleep any event that:
          ## * Follows an event marked as sleep, is a standing event and is between two sedentary events;
          ## * Follows an event marked as sleep, is a sedentary event and follows a sedentary event followed by a standing event;
          ## * Precedes an event marked as sleep, is a standing event and is between two sedentary events or
          ## * Precedes an event marked as sleep, is a standing event and precedes a standing event followed by a sedentary event.
          sleep = if_else(
            lag(sleep) & lag(ActivityType) == 0 & ActivityType == 1 & lead(ActivityType) == 0,
            TRUE,
            sleep
          ),
          sleep = if_else(
            lag(sleep) & lag(ActivityType, 2) == 0 & lag(ActivityType) == 1 & ActivityType == 0,
            TRUE,
            sleep
          ),
          sleep = if_else(
            lead(sleep) & lag(ActivityType) == 0 & ActivityType == 1 & lead(ActivityType) == 0,
            TRUE,
            sleep
          ),
          sleep = if_else(
            lead(sleep) & ActivityType == 0 & lead(ActivityType) == 1 & lead(ActivityType, 2) == 0,
            TRUE,
            sleep
          )
        ) %>%
        
        # Removing unnecessary variables
        select(-sleep_end, -sleep_start, -total_end, -total_start) %>%
        
        # Ungrouping
        ungroup()
      
      # Recording new sleep event count to be used to determine whether iteration should continue
      new <- sum(df$sleep, na.rm = TRUE)
    }
    
    # Once iteration has concluded...
    df %<>%
      
      # Grouping by reference day
      group_by(refday) %>%
      
      mutate(
        # Determining proportion of time in each reference day occupied by each activity type
        prop_sed = sum(Intervals[ActivityType == 0])/sum(Intervals),
        prop_stand = sum(Intervals[ActivityType == 1])/sum(Intervals),
        prop_step = sum(Intervals[ActivityType == 2])/sum(Intervals),
        
        # Rejecting steps during an event marked as sleep
        steps = if_else(
          sleep,
          0L,
          steps
        ),
        
        # Marking events not marked as sleep as waking wear with their duration as a new variable
        wwear = if_else(
          sleep,
          0,
          as.double(Intervals)
        ),
        
        # Summing steps and waking wear for each reference day
        sum_steps = sum(steps),
        sum_wwear = sum(wwear),
        
        # Invalidating days where:
        ## * Any one activity occupies more than 95% of the time;
        ## * Fewer than 500 valid steps were recorded or
        ## * The waking wear time was less than 10 hours.
        invalid_day = case_when(
          prop_sed >= 0.95 | prop_stand >= 0.95 | prop_step >= 0.95 ~ TRUE,
          sum_steps < 500 ~ TRUE,
          sum_wwear < 36000 ~ TRUE,
          TRUE ~ FALSE
        )
      ) %>%
      
      # Removing invalid days
      filter(!invalid_day) %>%
      
      mutate(
        # Recording the start of the first event in a series of sleep events as sleeptime
        sleeptime = if_else(
          sleep & !lag(sleep),
          Time,
          NA_POSIXct_
        ),
        
        # Recording the end of the last event in a series of sleep events as waketime
        waketime = if_else(
          sleep & !lead(sleep),
          Time + Intervals,
          NA_POSIXct_
        )
      ) %>%
      
      # Removing all events not marked as sleeptime or waketime
      filter(!is.na(sleeptime) | !is.na(waketime)) %>%
    
      # Retaining only reference day, sleeptime and waketime indicators
      select(refday, sleeptime, waketime) %>%
      
      # Filling NAs with sleeptime downwards and waketime upwards, such that each row represents one sleep spell with a sleeptime and a waketime
      fill(sleeptime, .direction = "down") %>%
      fill(waketime, .direction = "up") %>%
      
      # Removing duplicate rows so there is only one row per sleep spell
      distinct() %>%
      
      mutate(
        # Deriving duration of a sleep spell as waketime - sleeptime
        duration = abs(difftime(waketime, sleeptime, units = "secs")),
        
        # Summing sleep durations per reference day
        sleepdur = sum(duration)
      ) %>%
      
      # Removing unnecessary duration variable
      select(-duration) %>%
      
      # Ungrouping
      ungroup() %>%
      
      # Deriving median daily sleep duration if there are at least three values available, otherwise recording NA
      mutate(md_winkler_sleepdur = if_else(
        sum(!is.na(sleepdur)) >= 3,
        median(sleepdur, na.rm = TRUE),
        NA_real_
      )) %>%
      
      # Removing unnecessary variables
      select(-refday, -sleeptime, -waketime, -sleepdur) %>%
      
      # Removing duplicate rows so there is only one row left
      distinct() %>%
      
      # Pulling participant ID from the filename and adding it to the row of data
      mutate(ParticipantId = as.integer(str_sub(i, 1, 6)), .before = md_winkler_sleepdur)
    
    # Appending the participant's value to the running data frame
    assign(paste0("df", no),
           rbind(get(paste0("df", no)), df))
    
    # Removing working data frame
    rm(df)
  }
})

# Stopping cluster
stopCluster(cl)
```

## Combining and saving data
Finally, we can combine the parallel data frames to produce a single complete data frame, and write this to a .csv file.

```{r write}
# Binding data frames together
df <- reduce(mget(ls(pattern = "df")), rbind)

# Writing combined data frame to .csv file
fwrite(df, "winkler_sleep.csv")
```